---
title: "Turtle nesting at Thevenard Island"
author: "Kesia Savill (Murdoch University)"
date: "`r Sys.time()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
  pdf_document: default
---

```{r setup, include=FALSE, message=F}
knitr::opts_chunk$set(echo = TRUE)  # Show code
# knitr::opts_chunk$set(echo = FALSE) # Don't show code
library(wastdr)
library(dplyr)
library(magrittr)
library(lubridate)
library(leaflet)
library(mapview)
library(ggplot2)
wastdr::wastdr_setup()
# wastdr::wastdr_settings()
```

# Attribution
* Data sharing agreement with DBCA
* DBCA contacts
* Data point of truth is WA Sea Turtle Database 
  [WAStD](https://strandings.dpaw.wa.gov.au/), https://strandings.dpaw.wa.gov.au/ 

# Data

### Downloading the data from WAStD

```{r download_data, eval=F}
filter_to_end2016 <- . %>% dplyr::filter(
  date > dmy("17/11/2016"), 
  date < dmy("31/12/2016"))

filter_to_end2017 <- . %>% dplyr::filter(
  date > dmy("17/11/2017"), 
  date < dmy("31/12/2017"))

filter_tagging <- . %>% dplyr::filter(encounter_type == "tagging")

exclude_training <- . %>% dplyr::filter(site_id!=30, species!="corolla-corolla")

if (file.exists("data/tnejson.Rda")){
  load("data/tnejson.Rda")
} else {
  tnejson <- wastdr::get_wastd(
    'turtle-nest-encounters', 
    query = list(taxon = "Cheloniidae", 
                 format = "json",
                 area=17))
  save(tnejson, file = "data/tnejson.Rda")
}

tne <- parse_turtle_nest_encounters(tnejson)

tracks2016 <- tne %>% filter_to_end2016 %>% exclude_training
tracks2017 <- tne %>% filter_to_end2017 %>% exclude_training

save.image("data/data.RData")
```


Raw data contains all observations:
```{r resume_from_saved_data}
load("data/data.RData")
```

### Inspecting the data

```{r show_json}
# listviewer::jsonedit(tnejson$features)
```

Parsed data is tabular:
```{r show_dt}
DT::datatable(head(tne, n = 100))
```

### Further reading

Examples of analyses in R:

* [Example workbook](http://rpubs.com/florian_mayer/tracks)
* [wastdr homepage](https://parksandwildlife.github.io/wastdr/)
* [wastdr usage examples](http://rpubs.com/florian_mayer/turtles-wastdr)

# Map

```{r mapping_helpers}
species_colours <- tibble::tibble(
    species = c(
    "cheloniidae-fam",
    "chelonia-mydas",
    "eretmochelys-imbricata",
    "natator-depressus",
    "corolla-corolla",
    "lepidochelys-olivacea",
    "caretta-caretta"    
    ),
    species_colours = c(
    "gray",
    "green",
    "darkblue",
    "beige",
    "pink",
    "darkgreen",
    "orange"
    )
)

nest_type_text <- tibble::tibble(
    nest_type = c(
        "hatched-nest", 
        "successful-crawl",
        "track-not-assessed",
        "track-unsure",
        "nest",
        "false-crawl"),
    nest_type_text = c(
        "NH", 
        "T+N",
        "T+?",
        "N?",
        "N",
        "T")
)

add_lookups <- . %>% 
    dplyr::left_join(species_colours, by="species") %>%
    dplyr::left_join(nest_type_text, by="nest_type")

tracks_map <- function(track_data) {
  l <- leaflet::leaflet(width=800, height=600) %>% 
    addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
    addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
    clearBounds()
  
  tracks.df <-  track_data %>% split(track_data$species)
  
  names(tracks.df) %>%
    purrr::walk( function(df) {
      l <<- l %>%
        addAwesomeMarkers(
          data = tracks.df[[df]],
          lng = ~longitude, lat=~latitude,
          icon = leaflet::makeAwesomeIcon(
            text = ~nest_type_text,
            markerColor = ~species_colours),
          label=~paste(date, nest_age, species, nest_type),
          popup=~paste(date, nest_age, species, nest_type),
          group = df
        )
    })
  
  l %>%
    addLayersControl(
      baseGroups = c("Aerial", "Place names"),
      overlayGroups = names(tracks.df),
      options = layersControlOptions(collapsed = FALSE)
    )
}
```

### A simple map

```{r simple_map_tracks_2017}
simple_tracks_2017 <- tracks2017 %>% add_lookups
leaflet::leaflet(width=800, height=600) %>% 
    addProviderTiles("Esri.WorldImagery", group = "Aerial") %>%
    addProviderTiles("OpenStreetMap.Mapnik", group = "Place names") %>%
    clearBounds() %>%
    leaflet::addAwesomeMarkers(
          data = simple_tracks_2017,
          lng = ~longitude, lat=~latitude,
          icon = leaflet::makeAwesomeIcon(
            text = ~nest_type_text,
            markerColor = ~species_colours),
          label=~paste(date, nest_age, species, nest_type),
          popup=~paste(date, nest_age, species, nest_type),
          group = "Tracks 2017"
        ) %>%
    leaflet::addLayersControl(
      baseGroups = c("Aerial", "Place names"),
      overlayGroups = c("Tracks 2017"),
      options = layersControlOptions(collapsed = FALSE)
    )
```

### Tracks 2016

```{r map_tracks_2016}
tracks2016 %>% add_lookups %>% tracks_map
```

### Nests 2016

```{r maps_nests_2016}
all_nests_2016 <- tracks2016 %>% 
  add_lookups %>% 
  filter(nest_type %in% c("successful-crawl", "nest", "hatched-nest"))

fresh_nests_2016 <- tracks2016 %>% 
  add_lookups %>% 
  filter(nest_type == "successful-crawl")

fresh_nests_2016 %>% 
  tracks_map %T>% 
  mapshot(., file = "nests_fresh_2016.png")
```

### Tracks 2017

```{r maps_tracks_2017}
tracks2017 %>% add_lookups %>% tracks_map
```

### Nests 2017

```{r maps_nests_2017}
all_nests_2017 <- tracks2017 %>% 
  add_lookups %>% 
  filter(nest_type %in% c("successful-crawl", "nest", "hatched-nest"))

fresh_nests_2017 <- tracks2017 %>% 
  add_lookups %>% 
  filter(nest_type == "successful-crawl")

fresh_nests_2017 %>% 
  tracks_map %T>% 
  mapshot(., file = "nests_fresh_2017.png")
```

# Data summaries

### Helpers

```{r}
daily_species_by_type <- . %>% 
    filter(nest_age=="fresh") %>%
    group_by(date, species, nest_type) %>% 
    tally() %>%
    ungroup()

daily_summary <- . %>% 
    daily_species_by_type %>% 
    tidyr::spread(nest_type, n, fill=0) %>%
    DT::datatable(.)

tracks_ts <- function(data, title="Abundance"){
    data %>%
    daily_species_by_type %>% 
    {ggplot(data=., aes(x = date, y = n, colour = nest_type)) + 
            geom_point() + 
            # geom_smooth(method = "auto") +
            # geom_line() +
            ylim(0, NA) +
            scale_x_date(breaks = scales::pretty_breaks(),
                         labels = scales::date_format("%d %b %Y")) +
            labs(colour="Track/Nest Type") +
            xlab("Date") +
            ylab("Number counted per day") +
            ggtitle(title) +
            theme_light()}
}
```

### Tracks 2016

```{r track_summary_2016}
tracks2016 %>% 
  # filter(species=="natator-depressus") %>%
  daily_summary()
```


```{r track_plot_2016}
tracks2016 %>% 
  tracks_ts(title="All species: tracks and nests") %T>% 
  ggsave("data/2016_all_species_all_types.png", ., width=7, height=5, dpi=300)

tracks2016 %>% 
  filter(species=="natator-depressus") %>%
  tracks_ts(title="Flatbacks: tracks and nests") %T>% 
  ggsave("data/2016_flatbacks_all_types.png", ., width=7, height=5, dpi=300)

all_nests_2016 %>% 
  # filter(species=="natator-depressus") %>%
  tracks_ts(title="All species: fresh, incubating and hatched nests") %T>% 
  ggsave("data/2016_all_species_all_nests.png", ., width=7, height=5, dpi=300)

fresh_nests_2016 %>% 
  # filter(species=="natator-depressus") %>%
  tracks_ts(title="All species: only fresh nests") %T>% 
  ggsave("data/2016_all_species_fresh_nests.png", ., width=7, height=5, dpi=300)
```

### Tracks for 2017
```{r track_summary_2017}
tracks2017 %>% 
  # filter(species=="natator-depressus") %>%
  daily_summary()
```

```{r track_plot_2017}
tracks2017 %>% 
  tracks_ts(title="All species: tracks and nests") %T>% 
  ggsave("data/2017_all_species_all_types.png", ., width=7, height=5, dpi=300)

tracks2017 %>% 
  filter(species=="natator-depressus") %>%
  tracks_ts(title="Flatbacks: tracks and nests") %T>% 
  ggsave("data/2017_flatbacks_all_types.png", ., width=7, height=5, dpi=300)

all_nests_2017 %>% 
  # filter(species=="natator-depressus") %>%
  tracks_ts(title="All species: fresh, incubating and hatched nests") %T>% 
  ggsave("data/2017_all_species_all_nests.png", ., width=7, height=5, dpi=300)

fresh_nests_2017 %>% 
  # filter(species=="natator-depressus") %>%
  tracks_ts(title="All species: only fresh nests") %T>% 
  ggsave("data/2017_all_species_fresh_nests.png", ., width=7, height=5, dpi=300)
```
### Season summaries: helpers

```{r summary_helpers}
species_day_summary_tracks <- . %>% 
  # dplyr::mutate(
  #   month = lubridate::month(lubridate::ymd(date))
  # ) %>%
  dplyr::filter(nest_age == "fresh",
                nest_type %in% c("track-not-assessed", 
                                 "track-unsure", 
                                 "false-crawl", 
                                 "successful-crawl")) %>% 
  group_by(species, date) %>% 
  tally %>%
  dplyr::summarize("mean" = mean(n) %>% round(digits = 2),
                   "SE" = sd(n)/sqrt(length(n)), # %>% round(digits = 2),
                   "SD" = sd(n), # %>% round(digits = 2),
                   "min" = min(n),
                   "max" = max(n))

```

Caveat: The data are trimmed down to only the first half of the nesting season 
(represented by a field trip in November/December). There was a second field trip
in Jan/Feb capturing mostly hatched nests.

Caveat: The true min is unknown, as the data are presence only (unless surveys are used).
On days without surveys, there could have been fewer than the min observed tracks.

Caveat: The daily tracks stats above exclude the tagging data from the nesting 
area, and may include training data from the training area (accommodation).

Caveat: The data (both tracks and tags) show both succcessful and unsuccessful nesting attempts.

### Daily stats: tracks 2016

```{r tracks_daily_summary_2016}
tracks2016 %>% species_day_summary_tracks %>% knitr::kable(.)
```

### Daily stats: tracks 2017

```{r tracks_daily_summary_2017}
tracks2017 %>% species_day_summary_tracks %>% knitr::kable(.)
```

# Questions

## Has the population size changed between 2016 and 2017?
Data: Tracks and tags, 2016 and 2017, exclude training data, import tagging 2017 data.
Filter: all tracks and nesting records?

## Has the abundance of nesting turtles changed between 2016 and 2017?
Data: Tracks and tags, 2016 and 2017, exclude training data, import tagging 2017 data.
Filter: only fresh evidence (tracks: nest_age=="fresh", tags: all) 
of successful nesting (tracks: nest_type=="successful-crawl", 
tags: nesting=="present"), but this excludes incorrectly tracks with maybe a nest
(nest_type=="track-unsure") and tracks not assessed for nests 
(nest_type=="track-not-assessed"). This bias applies in the same way to both seasons.


The sampe sizes are:

* There were **`r nrow(tracks2016)`** tracks in 2016
* TODO fill in the rest
